---
title: "Cleaning and Variable Analysis"
header-includes:
  - \usepackage{float}
  - \usepackage{longtable}
  - \usepackage{array}
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)  # ggplot, data manipulation tools
library(gridExtra)  # arrange plots
library(MASS)       # extract residuals, do boxcox etc.
```

\section{Cleaning and Preparing the Data}
```{r}
dat = read.csv("train.csv", skip = 1)
validInd = sample(1:nrow(dat), nrow(dat)/4) 
train = dat[-validInd, ]
valid = dat[validInd, ]
trainX = train[ ,1:(ncol(dat)-1)]
trainY = train[ ,ncol(dat)]
validX = valid[ ,1:(ncol(dat)-1)]
validY = valid[ ,ncol(dat)]
```


\section{Univariate Analysis}


\subsection{Variable Descriptions}
\begin{table}[h!]
\centering
\footnotesize
\centering
\caption{Description of variables in the data set.}
\label{Table:VarDesc}
\renewcommand{\arraystretch}{1.2}
\begin{tabular}{| c || c | c | p{'10cm}|}
\hline
\textbf{Variable Name} & \textbf{Data Type} & \textbf{Role in model} &\textbf{Description}\\ \hline\hline
\texttt{default payment next month} & \texttt{Factor} & \texttt{Response} &1 = a default payment, 0 = no default\\ \hline
\texttt{LIMIT\_BAL} & \texttt{Numeric} & \texttt{Predictor} & Amount of credit of an individual, in NT dollars\\ \hline
\texttt{SEX} & \texttt{Factor} & \texttt{Predictor} & Sex of an individual; 1 = male, 2 = female\\ \hline
\texttt{EDUCATION} & \texttt{Factor} & \texttt{Predictor} & Education status of an individual; 1 = graduate school, 2 = university, 3 = high school, 4 = other education\\ \hline
\texttt{MARRIAGE} & \texttt{Factor} & \texttt{Predictor} & Martial status of an individual; 1 = married, 2 = single, 3 = other\\ \hline
\texttt{AGE} & \texttt{Numeric} & \texttt{Predictor} & Age of an individual\\ \hline
\texttt{PAY\_0} to \texttt{PAY\_6} & \texttt{Factor} & \texttt{Predictor} & History of payment of an individual, from April (\texttt{PAY\_6}) to September (\texttt{PAY\_0}) 2015; -1 = on time, other values are months of delay in repayment\\ \hline
\texttt{BILL\_AMT1} to \texttt{BILL\_AMT6} & \texttt{Numeric} & \texttt{Predictor} & Amount of bill statement, from April (\texttt{BILL\_AMT6}) to September (\texttt{BILL\_AMT1}) 2015, in NT dollars\\ \hline
\texttt{PAY\_AMT1} to \texttt{PAY\_AMT6} & \texttt{Numeric} & \texttt{Predictor} & Amount of previous payment, from April (\texttt{PAY\_AMT6}) to September (\texttt{PAY\_AMT1}) 2015, in NT dollars\\ \hline
\end{tabular}
\end{table}

\subsection{Univariate Plots}

```{r echo=FALSE, fig.height=10, fig.width=10}

p1 <- ggplot(dat, aes(x = default.payment.next.month)) +
  geom_histogram(bins = 2, col = "black") +
  ggtitle("Histogram of Default Payment Last Month (name?)") +
  xlab("Default Payment Last Month")

p2 <- ggplot(dat, aes(x = LIMIT_BAL)) +
  geom_histogram(bins = 50, col = "black") +
  ggtitle("Histogram of LIMIT_BAL") +
  xlab("Amount of credit of an individual")

p3 <- ggplot(dat, aes(x = SEX)) +
  geom_histogram(bins = 2, col = "black") +
  ggtitle("Histogram of Sex") +
  xlab("Sex of an individual; 1 = male, 2 = female")

p4 <- ggplot(dat, aes(x = EDUCATION)) +
  geom_histogram(bins = 4, col = "black") +
  ggtitle("Histogram of Education") +
  xlab("Education of an individual")

p5 <- ggplot(dat, aes(x = MARRIAGE)) +
  geom_histogram(bins = 4, col = "black") +
  ggtitle("Histogram of Marriage") +
  xlab("Martial status of an individual; 0 = ???, 1 = married, \n 2 = single, 3 = high school, 4 = other")

p6 <- ggplot(dat, aes(x = AGE)) +
  geom_histogram(binwidth = 1, col = "black") +
  ggtitle("Histogram of Age") +
  xlab("Age of an individual")

# Output plots in 3x2 grid
grid.arrange(p1, p2, p3, p4, p5, p6, nrow = 3)
```





