---
title: "Statistical Modelling III - Group Assignment"
header-includes:
  - \usepackage{float}
  - \usepackage{longtable}
  - \usepackage{array}
author: 
output:
  pdf_document: default
  html_document: default
  fig_caption: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Import libraries
library(tidyverse)
library(gridExtra)
library(caret)
library(MASS)
library(kableExtra)

setwd("E:/Documents/GitHub/SM3_Group_Project")
```

```{r include=FALSE}
# Read in the headers
headers <- read.csv(file = "train.csv", skip=1, header=F, nrows=1, as.is=T)

# Read in complete dataset
dat <- read.csv(file = "train.csv", skip = 2, header = F)

# Add the headers to dataset
colnames(dat) <- headers

# Remove the ID field
dat <- dat[ ,-1]

# Rename response variable for brevity and set as factor
names(dat)[ncol(dat)]<-"y"
dat$y = as.factor(dat$y)

# Split into training and test set
set.seed(250)
trainIndex <- createDataPartition(dat$y, p = .7, list = FALSE, times = 1)
train.set <- dat[trainIndex,]
test.set <- dat[-trainIndex,]
```

# 1. Introduction

For banks and other institutions offering credit card loans to customers, risk management and default detection have always been crucial to their business. Although in most jurisdictions there are legal consequences discouraging credit card default, it is still a key issue that lenders have to contend with. Predicting whether or not a given borrower is likely to default on their credit card debt is complex problem. Statistical models can an important play minimising the risk of lending to borrowers who then default on their debt. In order to reduce adverse financial impact of these kinds of events, lenders can use information about previous customers, including whether or not they have defaulted on their loans, to assess the likelihood that a potential new customer will default. 

Predictive modelling is the process of applying statistical techniques to reveal the relationship between a response variable and one or more predictor variables, inclulding whether a relationship exists at all. Depending on the problem, there are a great number statistical techniques that can be used to make predictions from data. Most techniques however, will fall into one of two categories, regression and classification. In regression models, the response variable is a continuous variable. In classificaiton models, the response  is a categorical variable. When the response can only take two possible values, the problem is one of *binary classification*. 

In this project, the data taken from the file train and we are provided 20,000 credit card customers consists of a single binary response, indicating default in the October 2005 payments and 23 predictor variables. We are going to demonstrate the process of deriving the best fit predictive model, which represents the relationship between default of credit card payment and some significant features. The task is to produce a function takes a n × 23 matrix of predictor variables as its input and produces a n-dimensional vector of predictions with 0 for no default and 1 for default as its output.

At the very first parts of this paper, we will present missing data handling and denoting, defective data cleaning, making sure variables are appropriately categorised into quantitative (continuous or discrete) or categorical. Histogram and bar-chart will be plotted to illustrate their distribution respectively. As the prior process of selecting model, bivariate analysis will be included continuous and categorical variables and conducted by producing box-plot and scatter-plot to examine the effect of predictors on defaulting, and the relationship will be stated clearly. The cross validation method also used to select the final model. The following section is the most significant where we derive the final model by comparing the best model that derivate by algorithms involve backwards, forwards, stepwise and automatic. And in our last sections, statistic summary and figures will be illustrated to support the assumptions checking of our final model. As verification of our model selection, a prediction will be made on the default of a customer by using the final model and some new data.


# 2. Data Description



# 3. Cleaning



### 3.1. Loading and Type Assignment




### 3.2. Missing Values




# 4. Description of Variables

\subsection{Variable Descriptions}
\begin{table}[h!]
\centering
\footnotesize
\centering
\caption{Description of variables in the data set.}
\label{Table:VarDesc}
\renewcommand{\arraystretch}{1.2}
\begin{tabular}{| c || c | c | p{'10cm}|}
\hline
\textbf{Variable Name} & \textbf{Data Type} & \textbf{Role in model} &\textbf{Description}\\ \hline\hline
\texttt{default payment next month} & \texttt{Factor} & \texttt{Response} &1 = a default payment, 0 = no default\\ \hline
\texttt{LIMIT\_BAL} & \texttt{Numeric} & \texttt{Predictor} & Amount of credit of an individual, in NT dollars\\ \hline
\texttt{SEX} & \texttt{Factor} & \texttt{Predictor} & Sex of an individual; 1 = male, 2 = female\\ \hline
\texttt{EDUCATION} & \texttt{Factor} & \texttt{Predictor} & Education status of an individual; 1 = graduate school, 2 = university, 3 = high school, 4 = other education\\ \hline
\texttt{MARRIAGE} & \texttt{Factor} & \texttt{Predictor} & Martial status of an individual; 1 = married, 2 = single, 3 = other\\ \hline
\texttt{AGE} & \texttt{Numeric} & \texttt{Predictor} & Age of an individual\\ \hline
\texttt{PAY\_0} to \texttt{PAY\_6} & \texttt{Factor} & \texttt{Predictor} & History of payment of an individual, from April (\texttt{PAY\_6}) to September (\texttt{PAY\_0}) 2015; -1 = on time, other values are months of delay in repayment\\ \hline
\texttt{BILL\_AMT1} to \texttt{BILL\_AMT6} & \texttt{Numeric} & \texttt{Predictor} & Amount of bill statement, from April (\texttt{BILL\_AMT6}) to September (\texttt{BILL\_AMT1}) 2015, in NT dollars\\ \hline
\texttt{PAY\_AMT1} to \texttt{PAY\_AMT6} & \texttt{Numeric} & \texttt{Predictor} & Amount of previous payment, from April (\texttt{PAY\_AMT6}) to September (\texttt{PAY\_AMT1}) 2015, in NT dollars\\ \hline
\end{tabular}
\end{table}

\subsection{Univariate Plots}

```{r echo=FALSE, fig.height=10, fig.width=10}

# p1 <- ggplot(dat, aes(x = y)) +
#   geom_histogram(bins = 2, col = "black") +
#   ggtitle("Histogram of Default Payment Last Month (name?)") +
#   xlab("Default Payment Last Month")

p2 <- ggplot(dat, aes(x = LIMIT_BAL)) +
  geom_histogram(bins = 50, col = "black") +
  ggtitle("Histogram of LIMIT_BAL") +
  xlab("Amount of credit of an individual")

p3 <- ggplot(dat, aes(x = SEX)) +
  geom_histogram(bins = 2, col = "black") +
  ggtitle("Histogram of Sex") +
  xlab("Sex of an individual; 1 = male, 2 = female")

p4 <- ggplot(dat, aes(x = EDUCATION)) +
  geom_histogram(bins = 4, col = "black") +
  ggtitle("Histogram of Education") +
  xlab("Education of an individual")

p5 <- ggplot(dat, aes(x = MARRIAGE)) +
  geom_histogram(bins = 4, col = "black") +
  ggtitle("Histogram of Marriage") +
  xlab("Martial status of an individual; 0 = ???, 1 = married, \n 2 = single, 3 = high school, 4 = other")

p6 <- ggplot(dat, aes(x = AGE)) +
  geom_histogram(binwidth = 1, col = "black") +
  ggtitle("Histogram of Age") +
  xlab("Age of an individual")

# Output plots in 3x2 grid
grid.arrange(p2, p3, p4, p5, p6, nrow = 3)
```





### 4.1. Continuous Variables




### 4.2. Catagorical Variables




# 5. Bivariate Analysis

```{r dat, echo=FALSE}
headers <- read.csv(file = "train.csv", skip = 1, header = F, nrows = 1, as.is = TRUE)
# Read in complete dataset
dat <- read.csv(file = "train.csv", skip = 2, header = FALSE)
# Add the headers to dataset
colnames(dat) <- headers
```

In order to determine the nature and strength of the relationship between each predictor variable and the response variable, default payment next month, plots of each of these relationships will be produced and investigated. Determing the relative associations of these predictor variables is important, as it allows us to consider the implication of their inclusion in the final model.

## 5.1 Continous Variables
We will begin by considering the relationship of the continuous predictor variables against default payment next month. These relationships will be analysed using side-by-side box plots.

## 5.1.1 Limiting Balance against default payment next month

Given limiting balance is the amount of credit available to an individual, in NT dollars, the main areas of interest will be whether having a higher or lower limiting balance will influence defaulting next month's payment. We will investigate this using side-by-side box plots, shown below.

```{r LIMT_BAL, echo=FALSE}
ggplot(aes(x=as.factor(`default payment next month`), y=LIMIT_BAL, group=`default payment next month`),data=dat) + geom_boxplot() + ggtitle("Relationship between Limiting Balance and \n default payment next month") + theme(plot.title = element_text(hjust = 0.5)) + ylab("Limiting Balance") + xlab("Default Payment Next Month")
```

From observing the box plots above, it is clear that both levels of default payment next month are positively skewed.
We also see that the medians for both levels lie marginally closer to the lower quartile value than the upper. Although the upper and lower quartile values for those who do not default next month's payment lie above those who do respectively, the interquartile range for both levels overlap. Consider the median values for both levels shown below.

```{r LIMT_BAL medians}
print(median(dat$LIMIT_BAL[dat$`default payment next month`==0]))
print(median(dat$LIMIT_BAL[dat$`default payment next month`==1]))
```

The median limiting balance for those who do not default next month's payment is 150000, and 90000 for those who do default next month's payment. Thus it could be suggested that having a larger limiting balance amount may decrease the likelihood of defaulting next month's payment, however there is not sufficiently strong evidence to definitively conlude this. It should also be noted that both levels have potential outlier candidates.

## 5.1.2 Age against default payment next month

The next variable to be considered is Age, another integer variable taking values between 21 and 75. The relationship between Age and default payment next month will be investigated through side-by-side box plots.

```{r AGE, echo=FALSE}
ggplot(aes(x=as.factor(`default payment next month`), y=AGE, group=`default payment next month`),data=dat) + geom_boxplot() + ggtitle("Relationship between Age and \n default payment next month") + theme(plot.title = element_text(hjust = 0.5)) + ylab("Age") + xlab("Default Payment Next Month")
```

From observing the side-by-side box plots, we notice that both are positively skewed. Furthermore, the spread and median values for both levels of default payment next month are very similar, suggesting that there is no apparent trend for defaulting next month's payment based on Age.

```{r AGE medians}
print(quantile(dat$AGE[dat$`default payment next month`==0]))
print(quantile(dat$AGE[dat$`default payment next month`==1]))
```

Upon further inspection, the median values and the lower quartile values are the same for both levels, begin 34 and 28 respectively. Additionally, the upper quartile values for those who do default next month's payment and those who do not are  42 and 41 respectively. This further supports the suggestion that age does not influence whether or not next month's payment is defaulted. There are also potential outlier candidates for both levels.

## 5.1.3 Pay_X against default payment next month
PAY_X is a measure of an individual's payment history, with:

1. PAY_6 corresponding to the individuals payment in April.

2. PAY_5 corresponding to the individuals payment in May.

3. PAY_4 corresponding to the individuals payment in June.

4. PAY_3 corresponding to the individuals payment in July.

5. PAY_2 corresponding to the individuals payment in August.

6. PAY_0 corresponding to the individuals payment in September.

Each PAY_X is an integer variable, taking values between -2 and 8 (corresponding to month's the payment of that month was overdue). Hence, the relationship between Pay_X and default payment next month will be investigated using side-by-side boxplots.


```{r PAY_x, echo=FALSE, fig.height=10, fig.width=10}
p1 <- ggplot(aes(x=as.factor(`default payment next month`), y=PAY_0, group=`default payment next month`),data=dat) + geom_boxplot() + ggtitle("Relationship between PAY_0 and \n default payment next month") + theme(plot.title = element_text(hjust = 0.5)) + ylab("PAY_0") + xlab("Default Payment Next Month")

p2 <- ggplot(aes(x=as.factor(`default payment next month`), y=PAY_2, group=`default payment next month`),data=dat) + geom_boxplot() + ggtitle("Relationship between PAY_2 and \n default payment next month") + theme(plot.title = element_text(hjust = 0.5)) + ylab("PAY_2") + xlab("Default Payment Next Month")

p3 <- ggplot(aes(x=as.factor(`default payment next month`), y=PAY_3, group=`default payment next month`),data=dat) + geom_boxplot() + ggtitle("Relationship between PAY_3 and \n default payment next month") + theme(plot.title = element_text(hjust = 0.5)) + ylab("PAY_3") + xlab("Default Payment Next Month")

p4 <- ggplot(aes(x=as.factor(`default payment next month`), y=PAY_4, group=`default payment next month`),data=dat) + geom_boxplot() + ggtitle("Relationship between PAY_4 and \n default payment next month") + theme(plot.title = element_text(hjust = 0.5)) + ylab("PAY_4") + xlab("Default Payment Next Month")

p5 <- ggplot(aes(x=as.factor(`default payment next month`), y=PAY_5, group=`default payment next month`),data=dat) + geom_boxplot() + ggtitle("Relationship between PAY_5 and \n default payment next month") + theme(plot.title = element_text(hjust = 0.5)) + ylab("PAY_5") + xlab("Default Payment Next Month")

p6 <- ggplot(aes(x=as.factor(`default payment next month`), y=PAY_6, group=`default payment next month`),data=dat) + geom_boxplot() + ggtitle("Relationship between PAY_6 and \n default payment next month") + theme(plot.title = element_text(hjust = 0.5)) + ylab("PAY_6") + xlab("Default Payment Next Month")

# Output plots in 3x2 grid
grid.arrange(p1, p2, p3, p4, p5, p6, nrow = 3)
```

Observing these plots, we see that they are all positively skewed, with relatively low median values for both factor levels. Furthermore, other than for PAY_0, the median for both levels are equal. However, the upper quartile for all plots, other than PAY_6, is larger for those who do default next month's payment. Intuitively this makes sense, as we would expect defaulting next month's payment to be more likely if previous month's payments have also been late. There are also outlier candidates present in each of the plots.

## 5.1.4 BILL_AMTX against default payment next month
BILL_AMTX is the amount appearing on the bill statement as follows:

1. BILL_AMT6 corresponding to the individuals statement in April.

2. BILL_AMT5 corresponding to the individuals statement in May.

3. BILL_AMT4 corresponding to the individuals statement in June.

4. BILL_AMT3 corresponding to the individuals statement in July.

5. BILL_AMT2 corresponding to the individuals statement in August.

6. BILL_AMT1 corresponding to the individuals statement in September.

As BILL_AMTX is a numeric variable, we will investigate its relationship with default payment next month using side-by-side boxplots.

```{r BILL_AMTX, echo=FALSE, fig.height=10, fig.width=10}
p1 <- ggplot(aes(x=as.factor(`default payment next month`), y=BILL_AMT1, group=`default payment next month`),data=dat) + geom_boxplot() + ggtitle("Relationship between BILL_AMT1 and \n default payment next month") + theme(plot.title = element_text(hjust = 0.5)) + ylab("BILL_AMT1") + xlab("Default Payment Next Month")

p2 <- ggplot(aes(x=as.factor(`default payment next month`), y=BILL_AMT2, group=`default payment next month`),data=dat) + geom_boxplot() + ggtitle("Relationship between BILL_AMT2 and \n default payment next month") + theme(plot.title = element_text(hjust = 0.5)) + ylab("BILL_AMT2") + xlab("Default Payment Next Month")

p3 <- ggplot(aes(x=as.factor(`default payment next month`), y=BILL_AMT3, group=`default payment next month`),data=dat) + geom_boxplot() + ggtitle("Relationship between BILL_AMT3 and \n default payment next month") + theme(plot.title = element_text(hjust = 0.5)) + ylab("BILL_AMT3") + xlab("Default Payment Next Month")

p4 <- ggplot(aes(x=as.factor(`default payment next month`), y=BILL_AMT4, group=`default payment next month`),data=dat) + geom_boxplot() + ggtitle("Relationship between BILL_AMT4 and \n default payment next month") + theme(plot.title = element_text(hjust = 0.5)) + ylab("BILL_AMT4") + xlab("Default Payment Next Month")

p5 <- ggplot(aes(x=as.factor(`default payment next month`), y=BILL_AMT5, group=`default payment next month`),data=dat) + geom_boxplot() + ggtitle("Relationship between BILL_AMT5 and \n default payment next month") + theme(plot.title = element_text(hjust = 0.5)) + ylab("BILL_AMT5") + xlab("Default Payment Next Month")

p6 <- ggplot(aes(x=as.factor(`default payment next month`), y=BILL_AMT6, group=`default payment next month`),data=dat) + geom_boxplot() + ggtitle("Relationship between BILL_AMT6 and \n default payment next month") + theme(plot.title = element_text(hjust = 0.5)) + ylab("BILL_AMT6") + xlab("Default Payment Next Month")

# Output plots in 3x2 grid
grid.arrange(p1, p2, p3, p4, p5, p6, nrow = 3)
```

We see that all of these plots are positively skewed. Additionally, all side-by-side boxplots show that the median values for both levels of default payment next month lie withtin very close proximity of each other. Furthermore, there is also close similarities with the interquartile spread in each plot.

```{r BILL_AMTX medians}
print(Med_Def0 <- c(median(dat$BILL_AMT1[dat$`default payment next month`==0]), median(dat$BILL_AMT2[dat$`default payment next month`==0]), median(dat$BILL_AMT3[dat$`default payment next month`==0]), median(dat$BILL_AMT4[dat$`default payment next month`==0]), median(dat$BILL_AMT5[dat$`default payment next month`==0]), median(dat$BILL_AMT6[dat$`default payment next month`==0])))

print(Med_Def1 <- c(median(dat$BILL_AMT1[dat$`default payment next month`==1]), median(dat$BILL_AMT2[dat$`default payment next month`==1]), median(dat$BILL_AMT3[dat$`default payment next month`==1]), median(dat$BILL_AMT4[dat$`default payment next month`==1]), median(dat$BILL_AMT5[dat$`default payment next month`==1]), median(dat$BILL_AMT6[dat$`default payment next month`==1])))

```

Inspecting the median values for the 6 plots shows that the median value for BILL_AMT1, BILL_AMT2 AND BILL_AMT3 are higher for those who do not default next month's payment, where as this is reversed for BILL_AMT4, BILL_AMT5 and BILL_AMT6. This suggests that having a higher bill amount in recent month's will, in general, reduce the likelihood of defaulting next month's payment. However, the overlap in interquartile spread suggests that this influence is minimal. There are once again outlier candidates for both levels of default payment next month.

## 5.1.5 PAY_AMTX against default payment next month
PAY_AMTX is the amount of the previous payment, with:

1. PAY_AMT6 corresponding to the payment made in April.

2. PAY_AMT5 corresponding to the payment made in May.

3. PAY_AMT4 corresponding to the payment made in June.

4. PAY_AMT3 corresponding to the payment made in July.

5. PAY_AMT2 corresponding to the payment made in August.

6. PAY_AMT1 corresponding to the payment made in September.

As PAY_AMTX is a numeric variable, we will investigate its relationship with default payment next month using side-by-side boxplots.

```{r PAY_AMTX, echo=FALSE}
p1 <- ggplot(aes(x=as.factor(`default payment next month`), y=PAY_AMT1, group=`default payment next month`),data=dat) + geom_boxplot() + ggtitle("Relationship between PAY_AMT1 and \n default payment next month") + theme(plot.title = element_text(hjust = 0.5)) + ylab("PAY_AMT1") + xlab("Default Payment Next Month")

p2 <- ggplot(aes(x=as.factor(`default payment next month`), y=PAY_AMT2, group=`default payment next month`),data=dat) + geom_boxplot() + ggtitle("Relationship between PAY_AMT2 and \n default payment next month") + theme(plot.title = element_text(hjust = 0.5)) + ylab("PAY_AMT2") + xlab("Default Payment Next Month")

p3 <- ggplot(aes(x=as.factor(`default payment next month`), y=PAY_AMT3, group=`default payment next month`),data=dat) + geom_boxplot() + ggtitle("Relationship between PAY_AMT3 and \n default payment next month") + theme(plot.title = element_text(hjust = 0.5)) + ylab("PAY_AMT3") + xlab("Default Payment Next Month")

p4 <- ggplot(aes(x=as.factor(`default payment next month`), y=PAY_AMT4, group=`default payment next month`),data=dat) + geom_boxplot() + ggtitle("Relationship between PAY_AMT4 and \n default payment next month") + theme(plot.title = element_text(hjust = 0.5)) + ylab("PAY_AMT4") + xlab("Default Payment Next Month")

p5 <- ggplot(aes(x=as.factor(`default payment next month`), y=PAY_AMT5, group=`default payment next month`),data=dat) + geom_boxplot() + ggtitle("Relationship between PAY_AMT5 and \n default payment next month") + theme(plot.title = element_text(hjust = 0.5)) + ylab("PAY_AMT5") + xlab("Default Payment Next Month")

p6 <- ggplot(aes(x=as.factor(`default payment next month`), y=PAY_AMT6, group=`default payment next month`),data=dat) + geom_boxplot() + ggtitle("Relationship between PAY_AMT6 and \n default payment next month") + theme(plot.title = element_text(hjust = 0.5)) + ylab("PAY_AMT6") + xlab("Default Payment Next Month")

# Output plots in 3x2 grid
grid.arrange(p1, p2, p3, p4, p5, p6, nrow = 3)
```

We see that all of these plots are positively skewed. Additionally, all side-by-side boxplots show that the median, lower quartile and upper quartile values lie within very close proximity. 

```{r PAY_AMTX medians}
print(Median_Def0 <- c(median(dat$PAY_AMT1[dat$`default payment next month`==0]), median(dat$PAY_AMT2[dat$`default payment next month`==0]), median(dat$PAY_AMT3[dat$`default payment next month`==0]), median(dat$PAY_AMT4[dat$`default payment next month`==0]), median(dat$PAY_AMT5[dat$`default payment next month`==0]), median(dat$PAY_AMT6[dat$`default payment next month`==0])))

print(Median_Def1 <- c(median(dat$PAY_AMT1[dat$`default payment next month`==1]), median(dat$PAY_AMT2[dat$`default payment next month`==1]), median(dat$PAY_AMT3[dat$`default payment next month`==1]), median(dat$PAY_AMT4[dat$`default payment next month`==1]), median(dat$PAY_AMT5[dat$`default payment next month`==1]), median(dat$PAY_AMT6[dat$`default payment next month`==1])))

```

Upon further inspection, the median values are marginally larger for those who do not default next month's payment, so there may be a very small influence. There are outlier candidates for both levels of default payment next month.

## 5.2 Categorical Variables
We now consider the relationship between the categorical predictor variables and the response variable. These will be explored using side-by-side box plots where appropriate, otherwise analysing the mean values.

## 5.2.1 Sex against default payment next month

The next variable to be considered is Sex. As Sex and default payment next month are both integer variables with two levels (0 and 1 for default payment next month, 1 and 2 for sex) their relationship will be investigated by considering the default rates for both males and females.

```{r SEX, echo=FALSE}
print(male_def <- sum(dat$`default payment next month`[dat$SEX==1])/sum(dat$SEX==1))
print(female_def <- sum(dat$`default payment next month`[dat$SEX==2])/sum(dat$SEX==2))
```

Considering the above output, approximately 29.10\% of males defaulted next month's payment, whereas only 25.23\% did. This suggests that males may be more likely, on average, to default next month's payment.

## 5.2.2 Education against default payment next month

Education is another integer variable, taking 7 different levels, ranging from 0 to 6. We will once again consider side-by-side boxplots to investigate their relationship.

```{r Education, echo=FALSE}
ggplot(aes(x=as.factor(`default payment next month`), y=EDUCATION, group=`default payment next month`),data=dat) + geom_boxplot() + ggtitle("Relationship between Education and \n default payment next month") + theme(plot.title = element_text(hjust = 0.5)) + ylab("Education") + xlab("Default Payment Next Month")
mean(dat$EDUCATION[dat$`default payment next month`==0])
mean(dat$EDUCATION[dat$`default payment next month`==1])
```

Not sure if we really need a plot for this one??

The mean level of education for those who do not default next month's payment is 1.84, whereas for those who do it's 1.89. Due to how close these are in value, it is unreasonable to suggest that the level of education of an individual will effect whether or not they default.

## 5.2.3 Marriage against default payment next month

We will now consider Marriage, another integer variable with four levels (0 up to 3). Because there are only four levels, no plot will be considered to investigate their relationship.

```{r MARRIAGE, echo=TRUE}
print(mean(dat$MARRIAGE[dat$`default payment next month`==0]))
print(mean(dat$MARRIAGE[dat$`default payment next month`==1]))
```

The mean marriage value for those who do not default next month's payment is 1.56, whereas this value is 1.53 for those who do. These values are once again so close in value that there is no suggestion of a relationship between these two variables.

## 5.3 Bivariate Analysis Between Predictors

### 5.1. Continuous Variables




### 5.2. Categorical Variables




### 5.3. Bivariate Analysis Between Predictors



# 6. Model Fitting

## Feature Selection

Choosing an appropriate subset from the available predcitor variables is an important step in selecting an accurate model. While the exclusion of important predictors will produce an incorrect model, adding too many redundant predictor terms can overcomplicate the model, making it harder to interpret without improving its accuracy. It can also reduce the statistical accuracy of parameter estimates. In machine learning, the process of selecting predictor variables for a model is know as *feature selection*. 

We perform feature selection on the credit dataset using the Recursive Feature Elimination, which is a backwards selection technique that begins with all available features, and iteratively eliminates the least important ones.    

```{r eval=FALSE, include=FALSE}
# Preprocess data first
# normalized <- preProcess(X)

# subsets <- (1:5)
# 
# set.seed(10)
# 
# ctrl <- rfeControl(functions = lmFuncs,
#                    method = "repeatedcv",
#                    repeats = 5,
#                    verbose = FALSE)
# 
# lmProfile <- rfe(X, y,
#                  sizes = subsets,
#                  rfeControl = ctrl)
# 
# lmProfile
```

# Model Selection

Given that this problem is one of binary classification, the following types of models were considered: 

* Logistic regression
* Support vector machines
* Decision tree
* Random forest
* AdaBoost

## Model Training

Each of the models considered is generated below using the`train()` function from the `caret` package. 

```{r echo=FALSE}
# Initialize cross validation control
fitControl <- trainControl(method = "repeatedcv", number = 10)

set.seed(489)

# Train logistic regression model
mdl_logistic <- train(y~., data=train.set, method="glm", family=binomial(), trControl=fitControl, preProcess=c("center", "scale"))

# Train linear SVM
# mdl_svm_lin <- train(y~., data=train.set, method = "svmLinear", trControl=fitControl, preProcess=c("center", "scale"), tuneLength = 10)

# Train non-linear SVM - requires tuning hyperparameter C
# mdl_svm_radial <- train(y~., data=train.set, method="svmRadial",
# trControl=fitControl, preProcess=c("center", "scale"), tuneLength = 10)

# # Train decision model
mdl_dt <- train(y~., data=train.set, method="ctree", trControl=fitControl)
# 
# # Train random forest
# mdl_rf <- train(y~., data=train.set, method="cforest",  trControl=fitControl)
# 
# # Train AdaBoost
# mdl_ada <- train(y~., data=train.set, method = "adaboost", trControl=fitControl)
```

## Feature Importance



## Model Testing and Evaluation

The trained models are tested against the training set below. Predicted values for the response variable are obtained, and then the accuracy of these predictions is stored using the `confusionMatrix()` function. 

```{r echo=FALSE}
# Generate predicted classifications on test set with each model
pred_logistic <- predict(mdl_logistic,test.set)
# pred_svm_lin <- predict(mdl_svm_lin,test.set)
# pred_svm_radial <- predict(mdl_svm_radial,test.set)
pred_dt <- predict(mdl_dt,test.set)
# pred_rf <- predict(mdl_rf,test.set)
# pred_ada <- predict(mdl_ada,test.set)

# Store the results
cm_logistic <- confusionMatrix(pred_logistic, test.set$y)
# cm_svm_lin_pred <-  confusionMatrix(svm_lin_pred,test.set$y)
# cm_svm_radial_pred <-  confusionMatrix(svm_radial_pred,test.set$y)
cm_dt_pred <-  confusionMatrix(pred_dt,test.set$y)
# cm_rf_pred <-  confusionMatrix(rf_pred,test.set$y)
# cm_ada_pred <-  confusionMatrix(ada_pred,test.set$y)

# Extract the accuracy of each model
accuracy_logistic <- cm_logistic$overall[1]
# accuracy_svm_lin <- confusionMatrix(svm_lin_pred,test.set$y)
# accuracy_svm_radial <-confusionMatrix(svm_radial_pred,test.set$y)
accuracy_dt <- cm_dt_pred$overall[1]
# accuracy_rf <- confusionMatrix(rf_pred,test.set$y)
# accuracy_ada <- confusionMatrix(ada_pred,test.set$y)

# Placeholders
accuracy_svm_lin <- 0
accuracy_svm_radial <- 0
accuracy_rf <- 0
accuracy_ada <- 0

models_list <- c("Logistic Regression", "Linear SVM", "RBF Kernel SVM", "Decision Tree", "Random Forest", "AdaBoost")
accuracy_list <- c(round(accuracy_logistic*100,3), round(accuracy_svm_lin*100,3), round(accuracy_svm_radial*100,3), round(accuracy_dt*100,3), round(accuracy_rf*100,3), round(accuracy_ada*100,3))

accuracy_tbl <- data.frame(Model=models_list, Accuracy=accuracy_list) 
```

`r kable(accuracy_tbl, digits = 2, caption = "Test Accuracy of Trained Models", format = "latex", booktabs = T)`



# 7. Final Model



# 8. Conclusion



# Reference


